# Template for running drug-disease gene expression profile similarity analysis
# From Butte et. al., Sci Transl Med 2011

```{r start time, echo=FALSE}
begin.time <- Sys.time()
```

## Initialization

Load required packages and source necessary functions.

```{r initialize, echo=TRUE}
library("GEOquery") # For pulling data from GEO
library("siggenes") # For running SAM analysis
library("affyPLM") # For rank normalization of ExpressionSets
library("cMap2data") # For drug signature ranked lists
source("butte_helper.R") # Helper functions for analysis
data(drugRL) # Load the drug ranked list data from cMap
```

## Read data
We read in the GEO IDs for the datasets we would like to use for our analysis.

```{r read in data, echo=TRUE}
geo_id <- read.csv("geo_ids.csv", header=T)
ids <- geo_id$GDS
```

Next, we get the list of up- and down-regulated genes via SAM analysis of each specified dataset.

```{r get gene lists, echo=TRUE}
glist <- sapply(ids,get_gene_list, simplify=F)
names(glist) <- ids
```

We now know which genes are up- and down-regulated, and rank them by fold change in expression.
Using this ranking, we compare each disease signature to each drug signature.

```{r calculate dds, echo=TRUE}
ddsmat <- matrix(0,dim(drugRL)[2],length(glist),dimnames=c(colnames(drugRL),names(glist))

for(i in 1:dim(drugRL)[2]){
	for(j in 1:length(glist)){
		ddsmat[i,j] <- calc_dds(drugRL[,i],glist[[j]])
	}
}
```
